---
title: "Higher Education"
author: "Kevin Otieno"
date: "`r Sys.Date()`"
output: html_document
---

## Higher Education

The data is from Kaggle repository, focus is to learn the higher education system in Pakistan, in terms of the number of university, distribution per province and classification the universities

```{r cars}
data <- read.csv('University_data.csv')


#Summary of the data
summary(data)

#University data

length(unique(data$University))
#Total number of universities 227
length(unique(data$Location))
#Locations where we have the universities

length(unique(data$Province))
#Province where we have the universities
length(unique(data$Campuses))
#Campuses accross
length(unique(data$Specialization))
#Areas of specialization
length(unique(data$Type))
#The university classification
#Establishment per year

yearly <- aggregate(Type ~ Established, data = data, FUN = length)

#Significant years
head(yearly[order(-yearly$Type),],10)
```
### Province visualization
The viz helps in learning on the distribution of the universities per province

* Punjab has more universities 

* Gilgit-Baltistan has the least number of universities

```{r}

library(ggplot2)
province <- aggregate(Type ~ Province, data = data, FUN = length)
head(province)

prov <- ggplot(province, aes(x = Province, y = Type, fill=Type))
prov + geom_bar(stat = 'identity', width = 0.4) +
  labs(title = "Number of Universities per Province",
       y ='Count of University') +
  theme(plot.title = element_text(size = 12,face = 'bold',hjust=0.5),
         axis.text.x = element_text(vjust = 0.6,size = 10, angle = 90),
        axis.title.x = element_text(vjust = -1))

```

### Specialization visualization

The viz helps in learning on the distribution of the universities per province

* General course is offered across most the university

* Agriculture & Veterinary, Art & design and Sufism has is prominent in the universities

```{r}
specialization <- aggregate(Type ~ Specialization, data = data, FUN = length)

spec <- ggplot(specialization, aes(x = Specialization, y = Type, fill = Type))
spec + geom_bar(stat='identity', width = 0.5) +
  labs(title = "Course Specialization",
       y ='Count of University') +
  theme(plot.title = element_text(size = 12,
                                  face = 'bold',
                                  hjust = 0.5),
        axis.text.x = element_text(vjust = 0.6,size = 10, angle = 90),
        axis.title.x = element_text(vjust = -1))
```

### Classification of the universities

* Most of the universities are public

* Military universities also feature

```{r}
type <- aggregate(Specialization ~ Type, data = data, FUN = length)
type

t <- ggplot(type, aes(x= Type, y = Specialization, fill= Specialization))
t + geom_bar(stat = 'identity', width = 0.5) +
  labs(title = 'University Classification',
       y = 'Count of specialization') +
  theme(plot.title = element_text(size = 12,
                                  face = 'bold',
                                  hjust = 0.5),
        axis.text.x = element_text(vjust = 0.6,size = 10))
```



